---
title: "Lab One- Josh Goldhaber"
name: "Josh Goldhaber" 
output: "pdf_document"
date: "2025-09-09"
---

```{r setup}
library(tidyverse)
library(ggplot2)
library(skimr)
```


```{r}
pizza <- readr::read_csv("pizza.csv")
```

3. Create a descriptive statistics table. Feel free to use any package or method you feel works best.

```{r}
skim_without_charts(pizza)
```


4. Produce a histogram on the temperature of pizza upon delivery. Is temperature skewed, if so, how?
```{r}

pizza |>
  ggplot( aes(x = temperature, )) +
    geom_histogram(binwidth = 3) +
    labs(
      x = "Pizza Temp",
      y = "Count",
      title = "Pizza Temperature",
      subtitle = "Binwidth = 3"
    ) +theme_minimal()

```
temperature is skewed left. 

5. Produce a barplot comparing the average time for delivery by branch location.

```{r}
pizza |>
  group_by(area) |>
  summarise(meanDeliveryTime = mean(delivery_min, na.rm = TRUE)) |>
  ggplot(aes(x = area, y = meanDeliveryTime, fill = area)) +
  geom_col() +
  labs(
    x = "Branch",
    y = "Average Delivery Time (mins)",
    title = "Average Delivery Time by Branch"
  ) +
  theme_minimal()

```


6. Create a boxplot to for delivery time comparing the times across the three branches.

```{r}
pizza |> 
  group_by(area)|>
  ggplot(aes(x = area, y = delivery_min, fill = area))+
  geom_boxplot() + 
  labs(
    x = "Branch",
    y = "Delivery Time",
    title = "Delivery Time per Branch",
    fill = "Areas" 
  )
```







